---
title: "Influenza A H7N9 in China, 2013"
format: dashboard
theme: lux
---

```{r}
# Load necessary packages
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")
pacman::p_load(
  tidyverse, outbreaks, bslib, bsicons, shiny, 
  rgeoboundaries, plotly, htmltools, reactable, sf
)
```

```{r}
# Load dataset
fluH7N9_china_2013 <- outbreaks::fluH7N9_china_2013

# Total number of recorded cases
total_cases <- nrow(fluH7N9_china_2013)

# Total number of cases that were hospitalized
total_cases_hospitalised <- fluH7N9_china_2013 %>%
  filter(!is.na(date_of_hospitalisation)) %>%
  nrow()

# Total number of deaths
total_death <- fluH7N9_china_2013 %>%
  filter(outcome == "Death") %>%
  nrow()

# Number of cases by province
no_cases_province <- fluH7N9_china_2013 %>%
  group_by(province) %>%
  summarise(no_of_cases = n(), .groups = 'drop')

# Plot: Number of Cases by Province
no_cases_province_plot <- no_cases_province %>%
  ggplot(aes(x = no_of_cases, y = reorder(province, no_of_cases))) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
    title = "Number of Cases by Province",
    x = "Number of Cases",
    y = "Province"
  ) +
  theme_minimal()

# Interactive plot
interactive_no_cases_province_plot <- ggplotly(no_cases_province_plot)
interactive_no_cases_province_plot

# Spatial distribution of cases by province
china_province <- gb_adm1("china")

# Join administrative regions with cases by province
province_case_data <- no_cases_province %>%
  full_join(china_province, by = c("province" = "shapeName")) %>%
  mutate(province = case_when(
    province == "Anhui" ~ "Anhui Province",
    province == "Beijing" ~ "Beijing Municipality",
    province == "Fujian" ~ "Fujian Province",
    province == "Hebei" ~ "Hebei Province",
    province == "Henan" ~ "Henan Province",
    province == "Hunan" ~ "Hunan Province",
    province == "Jiangsu" ~ "Jiangsu Province",
    province == "Jiangxi" ~ "Jiangxi Province",
    province == "Shangdong" ~ "Shandong Province",
    province == "Shanghai" ~ "Shanghai Municipality",
    province == "Taiwan" ~ "Taiwan Province",
    province == "Zhejiang" ~ "Zhejiang Province",
    TRUE ~ province
  ))

# Convert to spatial data frame
province_case_sf <- st_as_sf(province_case_data)

# Spatial distribution plot
spatial_distribution_plot <- ggplot(province_case_sf) +
  geom_sf(aes(fill = no_of_cases)) +
  scale_fill_viridis_c(option = "plasma", name = "Cases") +
  theme_minimal() +
  labs(
    title = "Spatial Distribution of Cases",
    subtitle = "Number of cases by province"
  )

# Render the spatial plot
plot(spatial_distribution_plot)

```


# HOME
## Row 1 {height=30%}


```{r}
value_box(
  title = "Total Cases",
  value = total_cases,
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", total_cases, " cases)"))
)

```


```{r}

value_box(
  title = "Hospitalized Cases",
  value = total_cases_hospitalised,
  showcase = bsicons::bs_icon("hospital"),
  theme = value_box_theme(bg = "#a4a832"),
  p(paste0("(", total_cases_hospitalised, " cases)"))
)
```

```{r}
value_box(
  title = "Deaths",
  value = total_death,
  showcase = bsicons::bs_icon("fire"),
  theme = value_box_theme(bg = "#a83232"),
  p(paste0("(", total_death, " deaths)"))
)

```


## Row 2 {height=75%}
### {width=25%}

```{r}
interactive_no_cases_province_plot

```


### {width=25%}

```{r}
plot(spatial_distribution_plot)
```


# DOWNLOAD DATA

```{r}
library(htmltools)

htmltools::browsable(
  tagList(
    reactable(fluH7N9_china_2013, 
              elementId = "2013_China_flu", 
              searchable = TRUE, 
              filterable = TRUE),
    tags$button("Download as CSV", 
                onclick = "Reactable.downloadDataCSV('2013_China_flu')")
  )
)

```

# ABOUT

This dashboard analyzes data on the 2013 influenza A H7N9 outbreak in China.

Key points:

- There were `r total_cases` total reported cases
- `r total_cases_hospitalised` cases were hospitalized  
- `r total_death` deaths occurred
- The most affected provinces were Zhejiang, Shanghai and Jiangsu

The data is sourced from a Dryad dataset by Kucharski et al (2014) who collated case details from various reports. This dashboard provides an interactive visualization of the outbreak's geographical distribution and key statistics.

Reference: 
Kucharski, A., Mills, H., Pinsent, A., Fraser, C., Kerkhove, M. V., Donnelly, C. A., & Riley, S. (2014). Distinguishing between reservoir exposure and human-to-human transmission for emerging pathogens using case onset data. *PLOS Currents Outbreaks*, 1. https://doi.org/10.1371/currents.outbreaks.e1473d9bfc99d080ca242139a06c455f
