---
title: "Influenza A H7N9 in China, 2013"
format: 
  dashboard:
    embed-resources: true
theme: lux
---

```{r}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, outbreaks, bslib, bsicons, shiny, 
               rgeoboundaries, plotly, htmltools, reactable,
               sf)

# view data
head(outbreaks::fluH7N9_china_2013)
```

```{r}
fluH7N9_china_2013 <- fluH7N9_china_2013

# total number of recorded cases

total_cases <- fluH7N9_china_2013 %>% 
  count(case_id) %>%   
  pull(n) %>%         
  summary()   

# Total number of cases that was hospitalised

total_cases_hospitalised <- fluH7N9_china_2013 %>% 
  filter(!is.na(date_of_hospitalisation)) %>%   
  count(date_of_hospitalisation) %>% 
  pull(n) %>% 
  summary()


total_death <- fluH7N9_china_2013 %>% 
  filter(outcome =="Death") %>% 
  count(outcome) %>% 
  pull(n) %>% 
  summary()   

# Number of cases by province
no_cases_province <- fluH7N9_china_2013 %>%
  group_by(province) %>%  
  summarise(no_of_cases = n())  

no_cases_province_plot <- no_cases_province %>% 
  ggplot(aes(x = no_of_cases, y = reorder(province, no_of_cases))) + 
  geom_bar(stat = "identity", fill = "skyblue") +  
  labs(title = "Number of Cases by Province",
       x = "Number of Cases",
       y = "Province") +
  theme_minimal()

# Show interactive plot by province
ggplotly(no_cases_province_plot)


# spatial distribution of cases by province

china_province <- gb_adm1("china")

# join adm regions to cases by province dataset

new_no_cases_join <- no_cases_province %>%
  left_join(china_province, c(province= "shapeName")) %>% 
  mutate(province=case_when(
    province== "Anhui"~ "Anhui Province",
    province== "Beijing" ~ "Beijing Municipality",
    province== "Fujian" ~ "Fujian Province",
    province== "Hebei" ~ "Hebei Province",
    province== "Henan" ~ "Henan Province",
    province== "Hunan"~ "Hunan Province",
    province== "Jiangsu"~ "Jiangsu Province",
    province== "Jiangxi"~ "Jiangxi Province",
    province== "Shangdong"~ "Shangdong Province",
    province== "Shangai"~ "Shangai Municipality",
    province== "Taiwan"~ "Taiwan Province",
    province== "Zhejiang"~ "Zhejiang Province"
    ))
 
library(ggplot2)
library(sf)
  
new_no_cases_final <- st_as_sf(new_no_cases_join)
class(new_no_cases_final)

new_no_cases_finalplot <- ggplot(new_no_cases_final) +  
  geom_sf(aes(fill = no_of_cases)) +  
  scale_fill_viridis_c(option = "plasma", name = "Cases") + 
  theme_minimal() +  
  labs(title = "Spatial Distribution of Cases",
       subtitle = "Number of cases by province")

  
plot(new_no_cases_finalplot)




```

# HOME 
## Row 1 {height=25%}

```{r}
value_box(
  title = "total_cases",
  value = highest_life_exp_value,
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", total_cases, ")"))
)
```

```{r}
value_box(
  title = "total_cases_hospitalised",
  value = highest_life_exp_value,
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", total_cases_hospitalised, ")"))
)
```

```{r}
value_box(
  title = "total_death",
  value = highest_life_exp_value,
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", total_death, ")"))
)
```

## Row 2 {height=75%}
### width {25%}
```{r}
ggplotly(new_no_cases_finalplot)
```

# DOWNLOAD DATA

```{r}

library(htmltools)

htmltools::browsable(
  tagList(
    reactable(fluH7N9_china_2013, 
              elementId = "2013_China_flu", 
              searchable = T, 
              filterable = T), 
    
tags$button("Download as CSV", 
            onclick = "Reactable.downloadDataCSV('gapminder-table')")
  )
)
```


# ABOUT

This dashboard analyzes data on the 2013 influenza A H7N9 outbreak in China.

Key points:

- There were `r total_cases` total reported cases
- `r total_hospitalized` cases were hospitalized  
- `r total_deaths` deaths occurred
- The most affected provinces were Zhejiang, Shanghai and Jiangsu

The data is sourced from a Dryad dataset by Kucharski et al (2014) who collated case details from various reports. This dashboard provides an interactive visualization of the outbreak's geographical distribution and key statistics.

Reference: 
Kucharski, A., Mills, H., Pinsent, A., Fraser, C., Kerkhove, M. V., Donnelly, C. A., & Riley, S. (2014). Distinguishing between reservoir exposure and human-to-human transmission for emerging pathogens using case onset data. *PLOS Currents Outbreaks*, 1. https://doi.org/10.1371/currents.outbreaks.e1473d9bfc99d080ca242139a06c455f
